A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Electron Beam Dynamics

### 2019-03-19 paper/submission/ZW10157-response-2.tex
small edits done?

Thank you for the additional comments.
We have moved all code implementation details to appendices.
In addition, further revisions have been made as suggested.
We hope to show we value the constructive criticism provided.
The details can be found in the following pages.

@@

My thanks, also, to the authors for updating the cited references.
When blocks of references are provided, conventionally, they are
listed chronologically to properly recognize the contributions of
pioneers, and the authors should reorder their citations to follow
this convention.

@@

III. A: paragraph 2
search algorithm to get stuck --> search algorithm getting
stuck

{\bf Our response:} {\color{blue} This grammar error is corrected in the updated manuscript.}

III. B. paragraph 3
checking for idle WORKER --> checking for idle WORKERs

{\bf Our response:} {\color{blue} This typo is corrected in the updated manuscript.}

Fig. 3 caption
$r_k$ --> $r_1$
The work $W_j$ are beam dynamics simulation within OPAL.
--> $W_j$ performs beam dynamics simulations within OPAL.

{\bf Our response:} {\color{blue} The figure caption has been corrected.}

III. C. paragraph 1
shown in Listing 2 --> shown in Listing 1

@@

general nature of our approach --> general nature of our
framework's software design

{\bf Our response:} {\color{blue} This suggestion was used on line xx.}

@@

{\bf Our response:} {\color{blue} This was corrected in the updated manuscript.}

@@

paragraph 3
As shown in Listing 5 --> As shown in Listing 3

IV. paragraph 1:
in Listing 3 --> in Listing 1

V. B. 2 paragraph 2:
, and the to prevent --> and to prevent

@@

paragraph 3
but from Fig. 9, but this is --> but from Fig. 9, this is

paragraph 3
15 less --> 15 fewer

V. B. 3
paragraph 2 (reported as corrected but is not in the version provided
to reviewers)
to to --> to

>>>>

Thank you for the additional comments.
We have moved all code implementation details to appendices.
In addition, further revisions have been made as suggested.
The details can be found in the following pages.

@@

When blocks of references are provided, conventionally, they are
listed chronologically to properly recognize the contributions of
pioneers, and the authors should reorder their citations to follow
this convention.

@@

III. A: paragraph 2
search algorithm to get stuck --$>$ search algorithm getting
stuck

III. B. paragraph 3
checking for idle WORKER --$>$ checking for idle WORKERs

Fig. 3 caption
$r_k$ --$>$ $r_1$
The work $W_j$ are beam dynamics simulation within OPAL.
--$>$ $W_j$ performs beam dynamics simulations within OPAL.

III. C. paragraph 1
shown in Listing 2 --$>$ shown in Listing 1

@@

general nature of our approach --$>$ general nature of our
framework's software design

@@

@@

paragraph 3
As shown in Listing 5 --$>$ As shown in Listing 3

IV. paragraph 1:
in Listing 3 --$>$ in Listing 1

V. B. 2 paragraph 2:
, and the to prevent --$>$ and to prevent

@@

paragraph 3
but from Fig. 9, but this is --$>$ but from Fig. 9, this is

paragraph 3
15 less --$>$ 15 fewer

{\bf Our response:} {\color{blue} All typos/corrections above should be incorporated into the updated manuscript.}

V. B. 3
paragraph 2 (reported as corrected but is not in the version provided
to reviewers)
to to --$>$ to

{\bf Our response:} {\color{blue} 
	We were unable to locate this typo. Can the reviewer please provide a line number?}

### 2019-03-19 paper/submission/prab.tex
small grammer and typo edits
small edits done?

All \textsc{Optimizer} implementations have to respect the API shown in
Listing 2.

@@

In most cases, forward solver implementations are simple wrappers to run
  an existing ``external'' simulation code using a set of design variables as
  input. In the case of the OPAL integration, the \texttt{main} function is
  playing the role of the ``forward solver''. To underline the general nature of our approach, 
  in a similar project, the described methods are used for cavity shape optimisation based on \cite{ARBENZ2008381}. 
As for the \textsc{Optimizer} component there exists a base class, labeled
  \texttt{Simulation} as common basis for all \textsc{Simulation}
  implementations.
In addition, this component also inherits from the \texttt{Worker} class,
  already implementing the polling protocol for default worker types.
As shown in the API in Listing 3, the \texttt{Worker} class expects an
  implementation to provide implementations for those three methods.

@@

We aimed at an easy and expressive way for users to specify multi-objective optimization problems.
Following the principle of keeping metadata (optimization and simulation input data) together, 
we decided to embed the optimization problem specification in the simulation input file by 
prefixing it with special characters, e.g. as annotations prefixed with a special character.
In some cases, it might not be possible to annotate the simulation input file.
By providing an extra input file parser, optimization problems can be read from stand-alone files.
To allow arbitrary constraints and objective expressions, such as
%
%{0.2cm}
\begin{Verbatim}[fontsize=\scriptsize]
  name: OBJECTIVE,
        EXPR="5 * average(42.0, "measurement.dat") + ENERGY";
\end{Verbatim}
%\vspace{0.2cm}
%
\noindent
An expression parser using Boost Spirit~\cite{boost} was implemented.
In addition to the parser, we need an evaluator able to evaluate an expression,
given a parse tree and variable assignments to an actual value.
Expressions arising in multi-objective optimization problems usually evaluate
to booleans or floating point values.
The parse tree, also denoted abstract syntax tree (AST), is constructed recursively while an expression is parsed.
Upon evaluation, all unknown variables are replaced with values, 
either obtained from simulation results or provided by other subtrees in the AST.
In this stage, the AST can be evaluated bottom-up and the desired result is
  returned after processing the root of the tree.

@@

All custom functions are registered with expression objects.
This is necessary to ensure that expressions know how they can resolve
  function calls in their AST.
As shown in Listing 5 this is done by creating a collection of Boost
  functions~\cite{boost3} corresponding to the
  available custom functions in expressions and passing this to the
  \textsc{Pilot}.

@@

The framework contains a wrapper implementing the API mentioned in
  Listing 3 for using \textsc{OPAL}~\cite{opal} as the forward solver.
\textsc{OPAL} provides different trackers for cyclotrons and linear
  accelerators with satisfactory parallel performance. 
With access to the \textsc{OPAL} forward solver, the framework is able to
  tackle a multitude of optimization problems arising in the domain of
  particle accelerators.
  The framework is also integrated into \textsc{OPAL} so that users can 
  define optimization problems within an input file, requiring no 
  additional knowledge or installation of the API to use it.

@@

\section{Example Optimization File}

>>>>

All \textsc{Optimizer} implementations have to respect the API shown in
Listing 1.

@@

In most cases, forward solver implementations are simple wrappers to run
  an existing ``external'' simulation code using a set of design variables as
  input. In the case of the OPAL integration, the \texttt{main} function is
  playing the role of the ``forward solver''. To underline the general nature of our framework's software design, 
  in a similar project, the described methods are used for cavity shape optimisation based on \cite{ARBENZ2008381}. 
As for the \textsc{Optimizer} component there exists a base class, labeled
  \texttt{Simulation} as common basis for all \textsc{Simulation}
  implementations.
In addition, this component also inherits from the \texttt{Worker} class,
  already implementing the polling protocol for default worker types.
As shown in the API in Listing 2, the \texttt{Worker} class expects an
  implementation to provide implementations for three methods.

@@

We aimed at an easy and expressive way for users to specify multi-objective optimization problems.
Following the principle of keeping metadata (optimization and simulation input data) together, 
we decided to embed the optimization problem specification in the simulation input file by 
prefixing it with special characters, e.g. as annotations prefixed with a special character.
In some cases, it might not be possible to annotate the simulation input file.
By providing an extra input file parser, optimization problems can be read from stand-alone files.
To allow arbitrary constraints and objective expressions, such as
%
%{0.2cm}
\begin{Verbatim}[fontsize=\scriptsize]
  name: OBJECTIVE,
        EXPR="5 * average(42.0, "measurement.dat") + ENERGY";
\end{Verbatim}
%\vspace{0.2cm}
%
\noindent
an expression parser using Boost Spirit~\cite{boost} was implemented.
In addition to the parser, we need an evaluator able to evaluate an expression,
given a parse tree and variable assignments to an actual value.
Expressions arising in multi-objective optimization problems usually evaluate
to booleans or floating point values.
The parse tree, also denoted abstract syntax tree (AST), is constructed recursively while an expression is parsed.
Upon evaluation, all unknown variables are replaced with values, 
either obtained from simulation results or provided by other subtrees in the AST.
In this stage, the AST can be evaluated bottom-up and the desired result is
  returned after processing the root of the tree.

@@

All custom functions are registered with expression objects.
This is necessary to ensure that expressions know how they can resolve
  function calls in their AST.
As shown in Listing 3 this is done by creating a collection of Boost
  functions~\cite{boost3} corresponding to the
  available custom functions in expressions and passing this to the
  \textsc{Pilot}.

@@

The framework contains a wrapper implementing the API mentioned in
  Listing 1 for using \textsc{OPAL}~\cite{opal} as the forward solver.
\textsc{OPAL} provides different trackers for cyclotrons and linear
  accelerators with satisfactory parallel performance. 
With access to the \textsc{OPAL} forward solver, the framework is able to
  tackle a multitude of optimization problems arising in the domain of
  particle accelerators.
  The framework is also integrated into \textsc{OPAL} so that users can 
  define optimization problems within an input file, requiring no 
  additional knowledge or installation of the API to use it.

@@

\section{Example Optimization File}\label{ap:inputfile}

### 2019-03-19 paper/submission/theor-experiment.tex
small edits done?

\subsubsection{Hyper parameter Scan}
While the optimization problem and goals were well defined (Subsection \ref{awaproblem}), 
it was not clear what the best hyper parameters for the genetic algorithm would be.
These parameters include gene mutation probability, mutation probability, 
recombination probability, number of individuals, 
and number of generations to complete. 
Given the beam line in Fig.~\ref{awa-linac},
four small optimization experiments were done with various hyper parameters. 
Similar to the time step scan, 
the goal of this exercise was to determine which set of optimization
parameters strongly influence the results, 
and whether there was a time to solution difference.
From here on, we will reference each experiment as ex-1, ex-2, ex-3, and ex-4
as shown in Table \ref{extable}. 
\begin{table}%[h!]
\begin{center}
\caption{Input Parameters for initial twenty four hour AWA optimization experiments. 
The gene mutation probability was equal to the mutation probability (not shown) in all four experiments. 
The max number of individuals per generation was~80.}
\label{extable}
\begin{tabular}{c|c|c|c}
\cline{1-4}
& Gene Mutation \qquad& Recombination \qquad & Number of completed \vspace{-0.75em}\\
Experiment & Probability & Probability &  generations \\ 
\cline{1-4}
ex-1 &  0.1  & 0.9  &  96 \\
ex-2 &  0.3  & 0.7  &  81 \\
ex-3 &  0.8  & 0.2  &  53 \\
ex-4 &  0.01 & 0.09 &  95 \\ 
\cline{1-4}
\end{tabular}
\end{center}
\end{table}

The maximum number of individuals per generation was fixed at 80. 
This number was chosen based on the node architecture, and the 
to prevent a prohibitive computational cost.  
Each experiment was allowed to run for twenty four hours, with 
a maximum generation limit of 100. 
We reduced the six objectives to four, 
and shortened the simulation time by moving the objectives further 
upstream to $s_1$ and $s_2$, the locations before and after the kicker, 
see Fig.~\ref{awa-tba}.  
The objectives include: $\varepsilon_{x}\left(s = s_1\right)\text{, } \varepsilon_{x}\left(s = s_2\right)$, $\text{rms}_{s}\left(s = s_1\right)\text{, and }  \text{rms}_{s}\left(s = s_2\right)$. 
The OPAL input file for these cases and all subsequent optimization runs in this paper, 
can be found at in following repository: \url{https://github.com/nneveu/awa-tba}.

After collection of the data for all four experiments, several metrics
were compared, including number of generations completed in twenty four hours and
Pareto fronts at $s_1$ and $s_2$, see Fig.~\ref{awa-linac}.
From Table \ref{extable}, we clearly see ex-3 is significantly 
slower, as it evaluated only 53 generations 
compared to the experiment with the maximum number, ex-1 at 96 generations.
Perhaps this trade off would be acceptable if the Pareto front was significantly 
improved, but from Fig. \ref{expareto}, but this is not the case.
Similar arguments can be made for ex-2, which evaluated about 15 less generations.
The Pareto fronts at $s_2$, are nearly identical. It is expected
this trend would continue given more time. 
When looking at the Pareto front at $s_1$, only ex-4 has a slightly 
larger range compared to the others.
With ex-2 and ex-3 eliminated due to evaluation time, 
and a slightly better Pareto front at $s_1$ for ex-4, 
the hyper parameters in ex-4 were chosen as the default values for subsequent runs.

>>>>

\subsubsection{Hyper parameter Scan}
While the optimization problem and goals were well defined (Subsection \ref{awaproblem}), 
it was not clear what the best hyper parameters for the genetic algorithm would be.
These parameters include gene mutation probability, mutation probability, 
recombination probability, number of individuals, 
and number of generations to complete. 
Given the beam line in Fig.~\ref{awa-linac},
four small optimization experiments were done with various hyper parameters. 
Similar to the time step scan, 
the goal of this exercise was to determine which set of optimization
parameters strongly influence the results, 
and whether there was a time to solution difference.
From here on, we will reference each experiment as ex-1, ex-2, ex-3, and ex-4
as shown in Table \ref{extable}. 
\begin{table}%[h!]
\begin{center}
\caption{Input Parameters for initial twenty-four hour AWA optimization experiments. 
The gene mutation probability was equal to the mutation probability (not shown) in all four experiments. 
The max number of individuals per generation was~80.}
\label{extable}
\begin{tabular}{c|c|c|c}
\cline{1-4}
& Gene Mutation \qquad& Recombination \qquad & Number of completed \vspace{-0.75em}\\
Experiment & Probability & Probability &  generations \\ 
\cline{1-4}
ex-1 &  0.1  & 0.9  &  96 \\
ex-2 &  0.3  & 0.7  &  81 \\
ex-3 &  0.8  & 0.2  &  53 \\
ex-4 &  0.01 & 0.09 &  95 \\ 
\cline{1-4}
\end{tabular}
\end{center}
\end{table}

The maximum number of individuals per generation was fixed at 80. 
This number was chosen based on the node architecture, and  
to prevent a prohibitive computational cost.  
Each experiment was allowed to run for twenty-four hours, with 
a maximum generation limit of 100. 
We reduced the six objectives to four, 
and shortened the simulation time by moving the objectives further 
upstream to $s_1$ and $s_2$, the locations before and after the kicker, 
see Fig.~\ref{awa-tba}.  
The objectives include: $\varepsilon_{x}\left(s = s_1\right)\text{, } \varepsilon_{x}\left(s = s_2\right)$, $\text{rms}_{s}\left(s = s_1\right)\text{, and }  \text{rms}_{s}\left(s = s_2\right)$. 
An example OPAL input file, used to perform the optimization in ex-1, 
can be found in Supplemental Material at [URL will be inserted by publisher].
All other optimization runs in the paper use this input file format.

After collection of the data for all four experiments, several metrics
were compared, including number of generations completed in twenty-four hours and
Pareto fronts at $s_1$ and $s_2$, see Fig.~\ref{awa-linac}.
From Table \ref{extable}, we clearly see ex-3 is significantly 
slower, as it evaluated only 53 generations 
compared to the experiment with the maximum number, ex-1 at 96 generations.
Perhaps this trade off would be acceptable if the Pareto front was significantly 
improved, but from Fig. \ref{expareto}, this is not the case.
Similar arguments can be made for ex-2, which evaluated about 15 fewer generations.
The Pareto fronts at $s_2$, are nearly identical. It is expected
this trend would continue given more time. 
When looking at the Pareto front at $s_1$, only ex-4 has a slightly 
larger range compared to the others.
With ex-2 and ex-3 eliminated due to evaluation time, 
and a slightly better Pareto front at $s_1$ for ex-4, 
the hyper parameters in ex-4 were chosen as the default values for subsequent runs.
